Predictability of Crash Modelling From the Data with Improved Quality Using Level Sets Surveillance System

### 2017-12-25 draft_new.Rmd
latex equation numbers
Add files via upload

---
title: ''
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```
#1. Introduction

#2. Evaluation Methodology
<describe how we evaluate data, the method, the process under different scenarios>

##2.1 Level sets

Level set methods are widely used in image segmentations and processing. One of its applciations is for pattern recognitions and shape reconstructions. Instead of continuous and multidimentional space, same implementation of level set method is used as a measure of pattern recognitions for discrete space crashes analysis in this study.

@@

<center> $\gamma = \{ x \in \Omega |  f(x) \ge c \}$ </center>

where $\gamma$ is the set of segments $x$'s that $f(x) \ge c$, $c$ is a threshold satisfying $\underset{c}{\operatorname{inf}} \frac{\| \gamma \|}{\| S_X \|} \le \alpha$ and $c$ can be determined as $\underset{x \in \gamma}{\operatorname{min}}f(x)$.

Some drawbacks of using level set methods in discrete space data are such as approximation of desired levels and solution to the ties with the same rates. Considering these two main issues in this study, level $\hat{\alpha}$ of sample is contraint on the proportion of highway segments that is closer to desired $\alpha$ but not greater than $\alpha$. Since same ranks represent that crashes are equally likely to happen to these tied segments, ties should be all inclusive or exclusive in our desired set as the level changes. These two issues are under control with two general approaches mentioned previously. The figure showed below demonstrates a comparison of crashes given different time periods under a level set method with $\alpha =$ 0.05 using highway data from Route 71, 2015. X-axis is the location ($x$ = Logmile) on this route whereas y-axis represents proportion of crashes $f(x)$. Threshold, $c$, (grey line) is set to be the minimal crash rate in the desired set $\gamma$. The true $\alpha$ levels for weekdays and weekend data are 0.0481 and 0.0377 with thresholds 0.0032 and 0.0039 repectively. 88 highway segments out of 1729 (or with a proportion of 0.0481) on route 71 with a crash rate equal or greater than 0.0032 approximately meets the desired level $\alpha =$ 0.05 for weekdays; similarily, 69 out of total 1729 segments on route 71 with a threshold equal or over 0.0039 satisfies this scenario. 
This setting is in a aspect of concertrating resources, say 0.05 in this example, and focusing on few highway segments with higher chance of having crashes with efficiency.

<center>
```{r}
load("~/all_plot.RData")
par(family = "serif", cex.axis = 0.7, mar = c(3.5, 2.5, 1, 1), las = 1,
    mfrow = c(1, 2), oma = c(2, 0, 0, 0))

plot.new()
plot.window(xlim = c(0, max(z$Logmile)), ylim = c(0,max(z$p)))
grid(nx = NULL, ny = NULL, col = "lightgray")
axis(1);axis(2)
segments(x0 = z1$Logmile, y0 = z1$p, y1 = 0)
points(x = z1[which(z1$p>=t1$thres), ]$Logmile, y = z1[which(z1$p>=t1$thres), ]$p, cex = 0.8)
abline(h = t1$thres, col = "darkgray")
box(which = "plot")
text(x = 32, y = t1$thres, labels = round(t1$thres, 4), pos = 3)
mtext("(a)", side = 1, line = 2, adj = 0.5)

plot.new()
plot.window(xlim = c(0, max(z$Logmile)), ylim = c(0,max(z$p)))
grid(nx = NULL, ny = NULL, col = "lightgray")
axis(1);axis(2)
segments(x0 = z2$Logmile, y0 = z2$p, y1 = 0)
points(x = z2[which(z2$p>=t2$thres), ]$Logmile, y = z2[which(z2$p>=t2$thres), ]$p, cex = 0.8)
abline(h = t2$thres, col = "darkgray")
box(which = "plot")
text(x = 32, y = t2$thres, labels = round(t2$thres, 4), pos = 3)
mtext("(b)", side = 1, line = 2, adj = 0.5)

fig.des <- expression(paste("Figure 1: weekdays (a) and weekend (b) with ", alpha, " = 0.05", sep = ""))
mtext(fig.des, side = 1, adj = 0.5, outer = TRUE)
```
</center>

https://link-springer-com.libdata.lib.ua.edu/book/10.1007%2F978-3-642-15352-5  
https://link-springer-com.libdata.lib.ua.edu/book/10.1007%2F978-3-319-01712-9

##2.2 Surveillance plots
Surveillance systems are built in respect of monitoring and capturing certain aberrations in the current or future process of tendency. These systems can be applied for detecting potential outbreak of epidemiological disease, and in essense used for evaluation of certain factors and profiles to acheive infection control in healthcare.

In conjuction with level set method, surveillance plots provide visualization of screening spatiotemporal highway data from distinct perspectives of monitoring crashes along highway segments. Under limited resources, say manpower, 10% or less of highway segments with the highest crash rates may be considered high priority and regularly kept under surveillance (fixed allocation based surveillance). Also, highway segments with a target percentage of crashes can also be determined and adjusted to capture different proportion of segments under various conditions (threshold based surveillance). Considering time influence, distributions of crashes can be expected either similiar or dissimilar. To observe and build up scenarios as described above, surveillance plots are essential to extend further exploratory and predictive analysis.

http://www.sciencedirect.com/science/article/pii/S187603411730206X  
http://www.sciencedirect.com/science/article/pii/S2093791116300610

###2.2.1 Fixed allocation based surveillance
The following is an example of fixed allocation based surveillance plots in two distinct way of presenting. Same data used for Figure 1 is also implemented in this example. Instead of single $\alpha$ level is observed in Figure 1, a sequence of $\alpha$ levels (from 0 to 1 with a increment of 0.01) are now concerned and showed in Figure 2. For (a) and (b), y-axis represents a cumulative proportion of crashes, and x-axis individually represents cumulative ordered segments size and top proportion of segments expected to be monitored, which is $\alpha$. The following paragraphs will describe this figures in detail.

<center>
```{r}
par(family = "serif", cex.axis = 0.7, mar = c(3.5, 2.5, 1, 1), las = 1,
    mfrow = c(1, 2), oma = c(2, 0, 0, 0))

plot.new()
plot.window(xlim = c(0, max(z$Logmile)), ylim = c(0, 1))
grid(nx = NULL, ny = NULL, col = "lightgray")
axis(1);axis(2)
lines(x = cumsum(z3$log.size), y = cumsum(z3$log.rate), type = "s")
lines(x = cumsum(z4$log.size), y = cumsum(z4$log.rate), type = "s", col = 4)
abline(v = h, col = "darkgrey")
box(which = "plot")
text(x = h, y = c(p1, p2, 0), labels = round(c(p1, p2, h), 4), pos = 2, col = c(1, 4, 1))
mtext("(a)", side = 1, line = 2, adj = 0.5)
legend("bottomright", c("Weekdays", "Weekend"), col = c(1, 4), lty = 1, bty = "n")

plot.new()
plot.window(xlim = c(0, 1), ylim = c(0, 1))
grid(nx = NULL, ny = NULL, col = "lightgray")
axis(1);axis(2)
lines(x = z1.lvl$alpha, y = z1.lvl$pct.events, type = "s")
lines(x = z2.lvl$alpha, y = z2.lvl$pct.events, type = "s", col = 4)
abline(v = 0.18, col = "darkgrey")
box(which = "plot")
text(x = thres, y = c(p3, p4, 0), labels = round(c(p3, p4, thres),4), pos = 2, col = c(1, 4, 1))
mtext("(b)", side = 1, line = 2, adj = 0.5)
legend("bottomright", c("Weekdays", "Weekend"), col = c(1, 4), lty = 1, bty = "n")

fig.des <- expression(paste("Figure 2: surveillance plots for weekdays and weekend under distinct perspectives", sep = ""))
mtext(fig.des, side = 1, adj = 0.5, outer = TRUE)
```
</center>

Table 1 is obtained with arrangement by order and shows the composition for Figure 2 (a). *diff* is the difference between each segments; *rank* is ranked by proportion of crashes in individual segments; *p* is proportion of crashes in individual segments; *Log.size* and *log.rate* are the sum of *diff* and *p* with the same ranked segments respectively (e.g. in *rank* = 7, 4 segments are tied, therefore with a *log.size* = *diff* $\times$ 4 = 0.0078 and *log.rate* = *p* $\times$ 4 = 0.0145, since the *diff* is all equal on route 71). It is plotted cumulative segmental differences versus cumulative proportion of crashes in ordered segments, and readible to see the surveillance under different desired lengths of highway.

Figure 2 (b), in analog of Figure 1 (a), will be showed simultaneously in the parentheses. The grey vertical line is the desired length of highway, top 6.26 *logmile* (or $\alpha =$ 0.18) being under surveillance. Under this scenario, 82.26% and 86.92% of total crashes will be primarily on top 6.26 *logmile* (or top 18% of total segments) for weekdays and weekend respectively.

```{r echo = FALSE, warning = FALSE, message = FALSE}
knitr::kable(head(z3, 10), caption = "Table 1: Partial output for Figure 2 (a)")
```

###2.2.2 Threshold based surveillance
Figure 3 shows an example of threshold based surveillance plot for weekdays and weekend. Y-axis is the proportion of segments above a given threshold whereas x-axis is desired threshold, which is the crash rate on a segment. Grey line is set to be a desired threshold 0.0019. 9.62% and 10.99% (or 176 and 201 segments) of total segments of their crash rates will be above a desired threshold of 0.0019 repectively. In comparison with fixed allocation based surveillance, it is also flexible to set up a desired threshold *c* instead of a desired proportion of segments $\alpha$ to be monitored.

@@

<center>
```{r}
par(family = "serif", cex.axis = 0.7, mar = c(3.5, 2.5, 1, 1), las = 1, oma = c(2, 0, 0, 0))

plot.new()
plot.window(xlim = rev(c(0, max(thr))), ylim = c(0, 1))
grid(nx = NULL, ny = NULL, col = "lightgray")
axis(1);axis(2)
lines(x = thr, y = z1.thres$size, type = "s")
lines(x = thr, y = z2.thres$size, type = "s", col = 4)
v <- 0.0019
abline(v = v, col = "darkgrey")
box(which = "plot")
p5 <- z1.thres[which(z1.thres$thr==v),]$size
p6 <- z2.thres[which(z2.thres$thr==v),]$size
text(x = v, y = c(p5+0.01, p6+0.04, 0), labels = round(c(p5, p6, v), 4), pos = 2, col = c(1, 4, 1))
legend("topleft", c("Weekdays", "Weekend"), col = c(1, 4), lty = 1, bty = "n")
mtext("Figure 3: threshold based surveillance for weekdays and weekend", 
      side = 1, adj = 0.5, outer = TRUE)
```
</center>

##2.3 Kolmogorov-Smirnov test
In analog of parametric method, *Q-Q* plots or *P-P* plots for assessing normality use quantile and percentile statistics to compare the magnitude of deviations between estimated and hypothetical distributions. Kolmogorov-Smirnov test (KS test) is a common nonparametric method to evaluate the goodness of fit between estimated and hypothetical distribution. For two-sample KS test, two random samples are compared to observe magnitude of differences based on their distributions. The primary approach of finding estimated distribution of a random variable is empirical cumulative distribution function (*ecdf*) or empirical distribution function (*edf*). An *ecdf* $S_n(x)$ is defined as:

<center> $S_n(x) = \sum_{i=1}^{n}{\textbf{I}(x_i \le x)}/n$ </center>

@@

<center> $D_n = \underset{x}{\operatorname{sup}} |S_n(x) - F_X(x)|$ </center>

@@

<center> $D_{m,n} = \underset{x}{\operatorname{max}} |S_m(x) - S_n(x)|$ </center>

If $D_{m,n} \ge d_0$, where $d_0 = c_{\alpha} \sqrt{\frac{1}{m} + \frac{1}{n}}$, $c_{\alpha}$ values can be found in any nonparametric materials, then $H_0$ is rejected, which implies $S_m(x)$ and $S_n(x)$ are from different distributions. Note that as $n, m \rightarrow \infty$, $c_{\alpha}$ is approximately $\sqrt{-\frac{1}{2} \operatorname{ln} (\frac{\alpha}{2})}$. Approximation of limiting $c_{\alpha}$ will be used for measure of distribution.

In this study, two-sample KS test is primarily conducted as a measure of association between distributions in analog of Least Significant Difference (LSD) for multiple comparisons in parametric methods. This test is contructed based on the snenarios used from a surveillance perspective whereas the surveillance is constructed based on the level set method; therefore, three ways of evaluations of highway data under different scenarios are seemingly distinct but closely tied together to organize a broaden perspecive of highway data structures.

Jean D. Gibbons, Subhabrata Chakraborti - Nonparametric Statistical Inference 4e (2003)

###2.3.1 Demonstration of conducting KS test
Recall from 2.2, surveillance using level set is comparing the proportion of crashes under various $\alpha$ values. Continuing using the same data as examples showed earlier, $S_m(x)$ and $S_n(x)$ are *cdf*'s for weekdays and weekend with same number of $\alpha$ values (*m* = *n* = 101) using level set method repectively. Aussume significance level is 0.05, then, $D_{m,n} =$ 0.085 ($d_0 \approx$ 0.1911) with *p-value* 0.96, and therefore, $H_0: S_m(x) = S_n(x)$ is rejected. This is suggested that two distributions are indistinguishable. It is possible that model based on weekdays might have similar predictability with model based on weekend. Instead of two, one model might be sufficient for prediction and surveillance.

The purpose of conducting a KS test on a fixed allocation based surveillance is that patterns of the top desired percentage of total segments for compared distributions are intrigued instead of the entire original highway patterns. In ordered segments, unnecessary segments can be easily discarded (beyond a $\alpha$ level), and the desired set (within $\alpha$ level) can be used to construct merely one model to predict weekdays and weekend patterns with better performance and precision. Using this procedure, model complexity and performance can both be ideally achieved.

##2.4 Jaccard Index plot
The Jaccard Index, or Jaccard similarity coefficient was introduced by Paul Jaccard and use as a measure of similarity of data. The Jaccard Index can be also extended to a Jaccard distance for measure of similarity in multidimensional space. General intersection-uion relationship will be applied in this paper. Suppose two samples are drawn, it is defined as a fraction of intersection and union sets of two samples.

<center> $J(A, B) = \frac{\| A \cap B \|}{\| A \cup B \|}$, where $A$ and $B$ represent two samples, $\|  \cdot \|$ is the size of a sample. </center>

Jaccard, Paul (1912), The distribution of the flora in the alpine zone

In accordance with level set method, Jaccard Index can be calculated under a sequence of $\alpha$ levels and the Jaccard Index plot is showed below in Figure 4. Grey lines are at $\alpha$ level 0.1 and 0.2, with percentages of matched segments 47.93% and 59.03% respectively for the Jaccard Index of weekdays and weekend sets. Note that percentage of matched rate stays constant starts from roughly $\alpha$ = 0.37, this is caused by no crash records within the rest of highway segments.

<center>
```{r}
par(family = "serif", cex.axis = 0.7, mar = c(3.5, 2.5, 1, 1), las = 1, oma = c(2, 0, 0, 0))
plot.new()
plot.window(xlim = c(0, 1), ylim = c(0, 1), ylab = "% matched segments")
grid(nx = NULL, ny = NULL, col = "lightgray")
axis(1);axis(2)
lines(x = alpha, y = m1, type = "s")
v1 <- c(0.1, 0.2)
abline(v = v1, col = "darkgrey")
box(which = "plot")
text(x = v1, y = m1[v1*length(alpha)+1]+0.02, 
     labels = round(m1[v1*length(alpha)+1], 4), pos = 2)
mtext("Figure 4: Jaccard Index for weekdays and weekend sets", 
      side = 1, adj = 0.5, outer = TRUE)
```
</center>

<PAI, etc>

#3. Predictability of Crash Data
In this section, tools mentioned in section 2 will be implemented as a surveillance process of quality of data. Process will follow 
<Focus on spatial prediction. Given a time period, how well can we say *where* crash occurs>

##3.1 Data sources
The dataset used is composed of two subsets: *Highway* data and *Crash* data, and focus on the state of Arkansas region. *Highway* data recorded the highway surface characteristics across the entire highway system in Arkansas. Categories and variables such as location  (logmile, longitude, latitude), traffic characteristics (average daily traffic), surface characteristics (type of road, sign of route, type of operation, median type and number of lanes) and others are included in this set. *Crash* data recorded individual crash cases at the spot. It includes time features (crash date, crash time), location (latitude, longitude, city, county, street), highway features (class of property, classification of trafficway, road system, intersection, junction relation), environmental features (light condition, road way surface condition) and crash conditions (interaction type, crash manner, number of fatals, number of injuries, number of vehicles) and others are included in this set.

@@

##3.2 Concentration of crashes: empirical performance measures

- Concentration of crashes: empirical performance measures (evaluate optimal performance - using only training data, no model)
		- vary segment
		- vary time (e.g., full year, each month, daytime, dow, etc)
		- show that e.g., mondays are more predictable than sundays. 
			- in extreme: for continuous time and space, then perfectly predictable. 
			- another extreme: worst (optimal) performance should be from using full year and large grid/segment size
	- use type 1 to predict, use type 2 to evaluate
		- type 1: day-time, type 2: night-time (etc, many scenarios)
		- show performance according the the evaluation methods we defined in 1.

>>>>

---
title: 'Predictability of Crash Modelling From the Data with Improved Quality Using Level Sets Surveillance System'
output: pdf_document
geometry: "left=2cm,right=2cm,top=2cm,bottom=2cm"
fig_caption: yes
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(dplyr)
library(bda)
library(knitr)
library(data.table)
library(kableExtra)
```

\section{1. Introduction}\label{sec1}
<build up scenario for problems, and summarize contents of section 2 and 3>

\section{2. Evaluation Methodology}\label{sec2}
Instead of going through traditional procedure of building statistical models such as negative-binomial and Poisson regression, couple useful concepts and tools of evaluating data before model building are introduced and modified to meet the need for the mothodology in this paper. These concepts include the level sets (\hyperref[sec21]{Section 2.1}), surveillance plots (\hyperref[sec22]{Section 2.2}), the Jaccard Index (\hyperref[sec23]{Section 2.3}) and Kolmogorov-Smirnov test (\hyperref[sec24]{Section 2.4}) and will be illustrated with a dataset from highway data in this order of comcepts.

\subsection{2.1 Level sets}\label{sec21}
Level sets methods are widely used in image segmentations and processing, one of its applciations is for pattern recognitions and shape reconstructions. Instead of continuous and multidimentional space, same implementation of level sets is used as a measure of pattern recognitions for discrete space crashes analysis in this study (\cite{ref1} and \cite{ref2}).

@@

\begin{equation}
  \gamma = \{ x \in \Omega |  f(x) \ge c \}
  \label{eq:1}
\end{equation}

where $\gamma$ is the set of segments $x$'s that $f(x) \ge c$, $c$ is a threshold that satisfies $\underset{c}{\operatorname{inf}} \frac{\| \gamma \|}{\| S_X \|} \le \alpha$ and $c$ can be determined as $\underset{x \in \gamma}{\operatorname{min}}f(x)$.

Some drawbacks of using level sets in discrete space data are such as approximation of desired levels and solution to the ties with the same rates. Considering these two main issues in this study, level $\hat{\alpha}$ of sample is contraint on the proportion of highway segments that is closer to desired $\alpha$ but not greater than $\alpha$. Since same ranks represent that crashes are equally likely to happen to these tied segments, ties should be all inclusive or exclusive in our desired set as the level changes. These two issues are under control with two general approaches mentioned previously. Figure \ref{fig:fig1} demonstrates a comparison of crashes given different time periods under level sets with $\alpha = 0.05$ using highway data from Route 71, 2015. X-axis is the location ($x$ = Logmile) on this route whereas y-axis represents proportion of crashes $f(x)$. Threshold, $c$, (gray line) is set to be the minimal crash rate in the desired set $\gamma$. The true $\alpha$ levels for weekdays and weekend data are 0.0481 and 0.0377 with thresholds 0.0032 and 0.0039 repectively. 88 highway segments out of 1729 (or with a proportion of 0.0481) on route 71 with a crash rate equal or greater than 0.0032 approximately meets the desired level $\alpha =$ 0.05 for weekdays; similarily, 69 out of total 1729 segments on route 71 with a threshold equal or over 0.0039 satisfies this scenario.
This setting is in a aspect of concertrating resources, say 0.05 in this example, and focusing on few highway segments with higher chance of having crashes with efficiency.

```{r, fig.align='center',fig.cap="\\label{fig:fig1}weekdays (a) and weekend (b) with $\\alpha$ = 0.05"}
load("wk_wkd.RData")
source_lines("wk_wkd.R", c(14:41))
```

\subsection{2.2 Surveillance plots}\label{sec22}
Surveillance systems are built in respect of monitoring and capturing certain aberrations in the current or future process of tendency. These systems can be applied for detecting potential outbreak of epidemiological disease, and in essense used for evaluation of certain factors and profiles to acheive infection control in healthcare (\cite{ref3} and \cite{ref4}).

In conjuction with level sets, surveillance plots provide visualization of screening spatiotemporal highway data from distinct perspectives of monitoring crashes along highway segments. Under limited resources, say manpower, 10% or less of highway segments with the highest crash rates may be considered high priority and regularly kept under surveillance (fixed allocation based surveillance). Also, highway segments with a target percentage of crashes can also be determined and adjusted to capture different proportion of segments under various conditions (threshold based surveillance). Considering time influence, distributions of crashes can be expected either similiar or dissimilar. To observe and build up scenarios as described above, surveillance plots are essential to extend further exploratory and predictive analysis.

\subsubsection{2.2.1 Fixed allocation based surveillance}\label{sec221}
The following is an example of fixed allocation based surveillance plots in two distinct way of presenting. Same data used for Figure \ref{fig:fig1} is also implemented in this example. Instead of single $\alpha$ level is observed in Figure \ref{fig:fig1}, a sequence of $\alpha$ levels (from 0 to 1 with a increment of 0.01) are now concerned and showed in Figure \ref{fig:fig2}. For (a) and (b), y-axis represents a cumulative proportion of crashes, and x-axis individually represents cumulative ordered segments size by crashes and top proportion of segments expected to be monitored, which is $\alpha$. The following paragraphs will describe this figure in detail.

```{r fig.align='center',fig.cap="\\label{fig:fig2}surveillance plots for weekdays and weekend under distinct perspectives $\\alpha$ levels"}
load("wk_wkd.RData")
source_lines("wk_wkd.R", c(54:99))
```

Table \ref{tab:tab1} is obtained with arrangement by order and shows the composition for \ref{fig:fig2} (a). *diff* is the difference between each segments; *rank* is ranked by proportion of crashes in individual segments; *p* is proportion of crashes in individual segments; *Log.size* and *log.rate* are the sum of *diff* and *p* with the same ranked segments respectively (e.g. in *rank* = 7, 4 segments are tied, therefore with a *log.size* = *diff* $\times$ 4 = 0.0078 and *log.rate* = *p* $\times$ 4 = 0.0145, since the *diff* is all equal on route 71). It is plotted cumulative segmental differences versus cumulative proportion of crashes in ordered segments, and readible to see the surveillance under different desired lengths of highway.

Figure \ref{fig:fig2} (b), in analog of Figure \ref{fig:fig2} (a), will be showed simultaneously in the parentheses. The gray vertical line is the desired length of highway, top 6.26 *Logmile* (or $\alpha =$ 0.18) being under surveillance. Under this scenario, 82.26% and 86.92% of total crashes will be primarily on top 6.26 *Logmile* (or top 18% of total segments) for weekdays and weekend respectively.

```{r}
kable(head(z3, 10), caption = "\\label{tab:tab1}Partial output for Figure 2.2 (a)")
```

\subsubsection{2.2.2 Threshold based surveillance}\label{sec222}
Figure \ref{fig:fig3} shows an example of threshold based surveillance plot for weekdays and weekend. Y-axis is the proportion of segments above a given threshold whereas x-axis is desired threshold, which is the crash rate on a segment. Gray line is set to be a desired threshold 0.0019. 9.62% and 10.99% (or 176 and 201 segments) of total segments of their crash rates will be above a desired threshold of 0.0019 repectively. In comparison with fixed allocation based surveillance, it is also flexible to set up a desired threshold *c* instead of a desired proportion of segments $\alpha$ to be monitored.

@@

```{r fig.align='center',fig.cap="\\label{fig:fig3}threshold based surveillance for weekdays and weekend"}
load("wk_wkd.RData")
source_lines("wk_wkd.R", c(108:129))
```

\subsection{2.3 Jaccard Index plot}\label{sec23}
The Jaccard Index, or Jaccard similarity coefficient was introduced by Paul Jaccard and use as a measure of similarity of data \cite{ref5}. The Jaccard Index can be also extended to a Jaccard distance for measure of similarity in multidimensional space. Suppose two samples are drawn, it is defined as a fraction of intersection and union sets of two samples.

\begin{equation}J(A, B) = \frac{\| A \cap B \|}{\| A \cup B \|}\label{eq:2}\end{equation}

where $A$ and $B$ represent two samples, $\|  \cdot \|$ is the size of a sample.

In this paper similar idea of Jaccard Index is used but in a conditional structure. It is defined as the following:

\begin{equation}J^{*}(A, B) = \frac{\| A \cap B \|}{\| A \|}\label{eq:3}\end{equation}

where $J^{*}(A, B)$ is conditioning on $A$. The diversity of sets $A$ and $B$ can be observed by comparing $J^{*}(A, B)$ and $J^{*}(B, A)$. The indexes should be close if both sets are similar, and vice versa.

Figure \ref{fig:fig4} shows the cumulative crash rate in original segments using (\ref{eq:3}). In \ref{fig:fig4} on the left, the black line is the exact cumulative crash rate on weekdays whereas the blue line is the cumulative crash rate with matched segments using reference weekdays, and right side shows the opposite condition.

```{r fig.align='center',fig.cap="\\label{fig:fig4}Comparison for weekdays and weekend using $J^{*}$ Index"}
load("wk_wkd.RData")
source_lines("wk_wkd.R", 144:167)
```

Based on the adjustment (\ref{eq:3}), $J^{*}$ can be calculated under a sequence of $\alpha$ levels is showed in Figure \ref{fig:fig5}. On the right of Figure \ref{fig:fig5}, under the same $\alpha$ level, the black line is the true cumulative crash rate in weekdays whereas the blue line shows the cumulative crash rate on weekend with the matched segments in weekdays, and the condition is opposite for the left side of Figure \ref{fig:fig5}.

```{r fig.align='center',fig.cap="\\label{fig:fig5}Comparison for weekdays and weekend under the level sets using $J^{*}$ Index"}
load("wk_wkd.RData")
source_lines("wk_wkd.R", 172:197)
```

Figure \ref{fig:fig4} and Figure \ref{fig:fig5} demonstrate the relationships of ordered distributions between weekdays and weekend under distinct aspects. Under the level sets setting, it is comparatively objective for comparisons for the reason that sizes of reference and subject are approximately equivalent. Great dissimilarities in distributions of ordered segments may be clearly showed in Figure \ref{fig:fig5}. In the next section, a metric to describe this similarity will be introduced and also served as a measure of quality of data.

\subsection{2.4 Kolmogorov-Smirnov test}\label{sec24}
In analog of parametric method, *Q-Q* plots or *P-P* plots for assessing normality use quantile and percentile statistics to compare the magnitude of deviations between estimated and hypothetical distributions. Kolmogorov-Smirnov test (KS test) is a common nonparametric method to evaluate the goodness of fit between estimated and hypothetical distribution (\cite{ref6}). For two-sample KS test, two random samples are compared to observe magnitude of differences based on their distributions. The primary approach of finding estimated distribution of a random variable is empirical cumulative distribution function (*ecdf*) or empirical distribution function (*edf*). An *ecdf* $S_n(x)$ is defined as:

\begin{equation}
  S_n(x) = \sum_{i=1}^{n}{\textbf{I}(x_i \le x)}/n
  \label{eq:4}
\end{equation}

@@

\begin{equation}
  D_n = \underset{x}{\operatorname{sup}} |S_n(x) - F_X(x)|
  \label{eq:5}
\end{equation}

@@

\begin{equation}
  D_{m,n} = \underset{x}{\operatorname{max}} |S_m(x) - S_n(x)|
  \label{eq:6}
\end{equation}

If $D_{m,n} \ge d_0$, where $d_0 = c_{\alpha} \sqrt{\frac{1}{m} + \frac{1}{n}}$, $c_{\alpha}$ values can be found in any nonparametric materials, then $H_0$ is rejected, which implies $S_m(x)$ and $S_n(x)$ are from different distributions. Note that as $n, m \rightarrow \infty$, $c_{\alpha}$ is approximately $\sqrt{-\frac{1}{2} \operatorname{ln} (\frac{\alpha}{2})}$. Approximation of limiting $c_{\alpha}$ will be used for measure of distributions.

In this study, two-sample KS test is primarily conducted as a measure of association between distributions in analog of Least Significant Difference (LSD) for multiple comparisons in parametric methods. Instead of calculating the *cdf* in using a raw data, it is done in a ranking way and *cdf* is not based on \ref{eq:4} but summing up the crash rates with the same rank. It is defined as follow:

\begin{equation}S^{*}_k(y) = \sum_{i=1}^{k}{p(Y \ge y_{(i)})}\label{eq:7}\end{equation}

where $S^{*}_k(y)$ is the cumulative crash rate, $0 \le S^{*}_k(y) \le 1$; *k* represents the rank; *Y* is the ordered segments and $y_{(i)}$ is the $i^{th}$ segments with the highest crash rate (ties are also considered). Since $y_{(1)}$ determines the maximum, $S^{*}_k(y)$ is showed as a way of survival function. (\ref{eq:7}) is viewed as foundation of the level sets applied in KS test.

A modified KS test using (\ref{eq:7}) is contructed based on the snenarios used from a surveillance perspective using the level sets; therefore, this is viewed as a solid metric to determine the diversity in distributions under different scenarios.

\subsubsection{2.4.1 Demonstration of conducting KS test}\label{sec241}
Recall from \hyperref[sec22]{Section 2.2}, surveillance using level sets is comparing the proportion of crashes under various $\alpha$ values. Continuing using the same data as examples showed earlier, $S^{*}_m(x)$ and $S^{*}_n(x)$ are *cdf*'s for weekdays and weekend with same number of $\alpha$ values (*m* = *n* = 101) using level sets repectively (see Figure \ref{fig:fig5}). Aussume significance level is 0.05, then, $D_{week} = 0.1631$ and $D_{weekend} = 0.2229$ ($d_0 \approx 0.1911$) with *p-value* 0.1364 and 0.0132 (Figure \ref{fig:fig6})

```{r fig.align='center',fig.cap="\\label{fig:fig6}KS test results for weekdays and weekend under the level sets using $J^{*}$ Index"}
load("wk_wkd.RData")
source_lines("wk_wkd.R", 203:229)
```

Therefore, it is significant in ordered distributions given reference weekend (Table \ref{tab:tab2}). This suggests that under the same size of segments, weekdays as reference to test weekend tends to be similar to true pattern. Also, weekdays as reference is more predictive than weekend as reference.

```{r}
kable(ks.out, row.names = TRUE, caption = "\\label{tab:tab2}KS results under the level sets in Figure \\ref{fig:fig5}")
```

The purpose of conducting a KS test on a fixed allocation based surveillance is that patterns of the top desired percentage of total segments for compared distributions are intrigued instead of the entire original highway patterns. In ordered segments, unnecessary segments can be easily discarded (beyond a $\alpha$ level), and the desired set (within $\alpha$ level) can be used as a predictive reference (say weekdays instead of weekend in the previous demonstration) to foresee patterns with better performance and precision.

\section{3. Predictability of Crash Data}\label{sec3}
In this section, the methods mentioned in previous section are implemented on a variety of scenarios, e.g. time variation and segments varation. These scenarios will be built under level sets, compared using modified Jaccard Index and evaluated with KS test. The visualization of prediction will be showed lastly.

\subsection{3.1 Data sources}\label{sec31}
The dataset used is composed of two subsets: *Highway* data and *Crash* data, and focus on the state of Arkansas region. *Highway* data recorded the highway surface characteristics across the entire highway system in Arkansas. Categories and variables such as location  (logmile, longitude, latitude), traffic characteristics (average daily traffic), surface characteristics (type of road, sign of route, type of operation, median type and number of lanes) and others are included in this set. *Crash* data recorded individual crash cases at the spot. It includes time features (crash date, crash time), location (latitude, longitude, city, county, street), highway features (class of property, classification of trafficway, road system, intersection, junction relation), environmental features (light condition, road way surface condition) and crash conditions (interaction type, crash manner, number of fatals, number of injuries, number of vehicles) and others are included in this set (\cite{ref7}).

@@

\subsection{3.2 Concentration of crashes: empirical performance measures}\label{sec32}
In previous demonstrations, crashes on the highway segments given different time periods, weekdays and weekend, are primarily showed a process of measuring data. KS test results are presented that weekdays may consider more predictive than weekend, i.e., a model can be built using weekdays set, and this model may also capture a similar pattern for weekend set. Other characteristrics of causes of crashes, e.g. time variation (\hyperref[sec321]{Section 3.2.1}) and segment variation (\hyperref[sec322]{Section 3.2.2}), will be compared through the same process as demonstrations earlier to uncover potential relationships in ordered distributions.

\subsubsection{3.2.1 Time variation}\label{sec321}
Given different time periods, distributions of crashes are varied. Several comparisons of distributions from time to time will be illustrated at the various significance level. The reason to the various significance level is that crashes distributed on an entire route are rare and more sparse with the consideration of time. Small change in distribution results in tremendous or  negligible extremes. Therefore, the choice of significance level is arbituary to an observable extent of relationships in time variation.

Note that the approximation of KS test and rare events condition result in a inconsistent in maximal distance and in *p-value*, so the magnitude of differences for both measures in distributions for months of year can be seen in Table \ref{tab:tab3} and \ref{tab:tab4}. The month used as reference is showed on the row whereas the compared groups are showed on the column. Cells in the table highlights in red to show the behavior under the significance level of 0.3. If only consider maximal distance in distributions ($D_{0.3} = 0.1722$ for this setting), May, June and  July (see from row vectors) which only have one significant pair under level sets individually, are appear to have better performance than any other months as reference.

```{r}
load("moy.RData")
data.frame(ks1$max.d) %>% 
  mutate(jan = cell_spec(jan, "latex", color = ifelse(jan > ks1$bound, "red", "black")),
         feb = cell_spec(feb, "latex", color = ifelse(feb > ks1$bound, "red", "black")),
         mar = cell_spec(mar, "latex", color = ifelse(mar > ks1$bound, "red", "black")),
         apr = cell_spec(apr, "latex", color = ifelse(apr > ks1$bound, "red", "black")),
         may = cell_spec(may, "latex", color = ifelse(may > ks1$bound, "red", "black")),
         jun = cell_spec(jun, "latex", color = ifelse(jun > ks1$bound, "red", "black")),
         jul = cell_spec(jul, "latex", color = ifelse(jul > ks1$bound, "red", "black")),
         aug = cell_spec(aug, "latex", color = ifelse(aug > ks1$bound, "red", "black")),
         sep = cell_spec(sep, "latex", color = ifelse(sep > ks1$bound, "red", "black")),
         oct = cell_spec(oct, "latex", color = ifelse(oct > ks1$bound, "red", "black")),
         nov = cell_spec(nov, "latex", color = ifelse(nov > ks1$bound, "red", "black")),
         dec = cell_spec(dec, "latex", color = ifelse(dec > ks1$bound, "red", "black"))) %>% 
kable("latex", row.names = TRUE, caption = "\\label{tab:tab3}KS test: max.d using level sets for months of year with significance level = 0.3", escape = FALSE, booktabs = TRUE, linesep = "")
```
